\documentclass[a4paper,11pt]{article}

\usepackage{psychtools}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{parskip}

\hypersetup{
  colorlinks=true,
  urlcolor=blue,
  linkcolor=black,
}

\newcommand{\pkg}[1]{\textsf{#1}}
\newcommand{\cmd}[1]{\texttt{\textbackslash #1}}
\newcommand{\opt}[1]{\texttt{#1}}
\newcommand{\meta}[1]{\ensuremath{\langle}\textit{#1}\ensuremath{\rangle}}

\title{\pkg{psychtools} --- LaTeX Macros for Psychology Manuscripts\\[0.5ex]
  \large Version 1.0}
\author{Ansgar Endress\\\texttt{ansgar.endress@gmx.net}}
\date{2026/06/26}

\begin{document}
\maketitle
\tableofcontents
\bigskip\hrule\bigskip

%% ---------------------------------------------------------------
\section{Introduction}

\pkg{psychtools} provides a collection of convenience macros for writing
empirical psychology and cognitive science manuscripts in LaTeX. It addresses
four common needs:

\begin{itemize}
  \item Compact, italicised notation for standard test statistics and
        effect sizes (\S\ref{sec:stats}).
  \item Supplementary-materials appendices with \texttt{S}-prefixed numbering
        and an optional TOC (\S\ref{sec:appendix}).
  \item Draft editing aids: conditional footnotes and section headers that
        can be toggled off for the final submission (\S\ref{sec:draft}).
  \item Citation shortcuts for \pkg{apacite} users (\S\ref{sec:cite}).
\end{itemize}

\noindent Load the package in the usual way:
\begin{verbatim}
  \usepackage{psychtools}
  % or with options:
  \usepackage[excludedetails,appendixtoc]{psychtools}
\end{verbatim}

\pkg{psychtools} requires \pkg{tocloft}, \pkg{xcolor}, \pkg{subcaption},
\pkg{xspace}, and \pkg{tikz}, all of which are part of standard TeX~Live and
MiK\TeX{} distributions.

%% ---------------------------------------------------------------
\section{Package Options}

\begin{center}
\begin{tabular}{lll}
\toprule
Option & Default & Effect \\
\midrule
\opt{includedetails} & on & Show draft footnotes and section headers \\
\opt{excludedetails} & ---& Hide all draft content (use for submission) \\
\opt{appendixtoc}   & off & Add a TOC for the supplementary materials \\
\opt{nocolorize}    & --- & Suppress red colouring of draft content \\
\opt{nocitecommands}& --- & Disable the \pkg{apacite} citation helpers \\
\bottomrule
\end{tabular}
\end{center}

%% ---------------------------------------------------------------
\section{Statistical Notation}
\label{sec:stats}

The commands in Table~\ref{tab:stats} typeset common test statistics and
effect sizes in italic math style, matching APA formatting requirements.
Each command includes a trailing \cmd{xspace} so it can be used freely in
running text without extra braces.

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
Command & Output & Meaning \\
\midrule
\cmd{T}  & \T  & Student's $t$ \\
\cmd{F}  & \F  & $F$ ratio \\
\cmd{Z}  & \Z  & $Z$ score \\
\cmd{p}  & \p  & $p$ value \\
\cmd{M}  & \M  & Mean \\
\cmd{SD} & \SD & Standard deviation \\
\cmd{SE} & \SE & Standard error \\
\cmd{D}  & \D  & Cohen's $d$ \\
\cmd{CI} & \CI & 95\% confidence interval \\
\cmd{et} & \et & Eta-squared \\
\cmd{etp}& \etp& Partial eta-squared \\
\cmd{U}  & \U  & Mann--Whitney $U$ \\
\cmd{W}  & \W  & Wilcoxon $W$ \\
\bottomrule
\end{tabular}
\caption{Statistical notation commands.}
\label{tab:stats}
\end{table}

\paragraph{Example}
\begin{verbatim}
  The effect was significant, \T(38) = 2.45, \p = .019, \D = 0.77.
\end{verbatim}
Produces: The effect was significant, \T(38) = 2.45, \p = .019, \D = 0.77.

\paragraph{Compatibility warning}
\label{sec:compat}
The short command names (\cmd{T}, \cmd{F}, \cmd{Z}, \cmd{p}, \cmd{M},
\cmd{U}, \cmd{W}) may conflict with commands defined by math or font
packages. To minimise the risk:
\begin{itemize}
  \item Load \pkg{psychtools} \emph{before} any math-heavy packages.
  \item If you encounter redefinition errors, check whether another
        loaded package defines the conflicting name and either remove
        that package or replace it with one that does not.
\end{itemize}

%% ---------------------------------------------------------------
\section{Supplementary Materials Appendix}
\label{sec:appendix}

Many psychology journals require supplementary materials numbered
\texttt{S1}, \texttt{S2},~\ldots{} with figures and tables prefixed
\texttt{S1}, \texttt{S2},~\ldots{} as well. The commands below set
up this numbering automatically.

\subsection{\cmd{myappendix}}

Place \cmd{myappendix} at the point in your document where the
supplementary materials begin. It:

\begin{itemize}
  \item resets the page counter to \texttt{S1};
  \item resets section, figure, and table counters with the \texttt{S}
        prefix;
  \item prints the appendix title (from \cmd{appendixname});
  \item optionally inserts a TOC for the supplementary materials
        (if the \opt{appendixtoc} option is active).
\end{itemize}

You should set \cmd{appendixname} in your document before calling
\cmd{myappendix}:

\begin{verbatim}
  \renewcommand{\appendixname}{Supplementary Online Materials}
  \myappendix
\end{verbatim}

If \cmd{appendixname} is not set, the prefix defaults to \texttt{S}.
The following values of \cmd{appendixname} produce a two-letter prefix:

\begin{center}
\begin{tabular}{ll}
\toprule
\cmd{appendixname} & Prefix \\
\midrule
\texttt{Supplementary Information} & \texttt{SI} \\
\texttt{Supplementary Materials} & \texttt{SM} \\
\texttt{Supplementary Online Materials} & \texttt{SM} \\
anything else & \texttt{S} \\
\bottomrule
\end{tabular}
\end{center}

\subsection{\cmd{appsection}\{\meta{title}\}}

Use \cmd{appsection} for sections inside the supplementary materials.
From the second section onwards it inserts a \cmd{clearpage} before
the new section, keeping the layout clean.

\begin{verbatim}
  \myappendix
  \appsection{Stimuli}
  ...
  \appsection{Additional Analyses}
  ...
\end{verbatim}

\subsection{Appendix TOC}

With the \opt{appendixtoc} option, \pkg{psychtools} suppresses the
main document TOC from showing supplementary-materials entries
(by setting \texttt{tocdepth} to $-2$ globally) and then restores
it inside \cmd{myappendix}, where a separate local TOC is printed.

%% ---------------------------------------------------------------
\section{Draft Editing Aids}
\label{sec:draft}

These commands are visible when the \opt{includedetails} option is
active (the default) and invisible when \opt{excludedetails} is used.
Draft content is coloured red (unless \opt{nocolorize} is passed).

\begin{center}
\begin{tabular}{lp{8.5cm}}
\toprule
Command & Effect \\
\midrule
\cmd{footnotewithdetails}\{\meta{text}\} &
  Inserts a red footnote marked ``(Footnote will be removed)''. \\
\cmd{draftsection}\{\meta{title}\} &
  Inserts a red section header marked ``(Header will be removed)''. \\
\cmd{draftsubsection}\{\meta{title}\} &
  Same for subsections. \\
\cmd{draftsubsubsection}\{\meta{title}\} &
  Same for subsubsections. \\
\cmd{colorize}\{\meta{text}\} &
  Colours \meta{text} red; passes through unchanged with \opt{nocolorize}. \\
\bottomrule
\end{tabular}
\end{center}

%% ---------------------------------------------------------------
\section{Citation Helpers (apacite)}
\label{sec:cite}

These commands are active when \pkg{apacite} is loaded and the
\opt{nocitecommands} option is not given.

\begin{center}
\begin{tabular}{ll}
\toprule
Command & Output \\
\midrule
\cmd{citeeg}\{\meta{key}\} & (e.g., Author Year) \\
\cmd{cites}\{\meta{key}\} & Author's Year \\
\bottomrule
\end{tabular}
\end{center}

If \pkg{apacite} is not loaded, basic fallbacks using \cmd{cite} and
\cmd{citeauthor}/\cmd{citeyear} are provided with a package warning.

%% ---------------------------------------------------------------
\section{Miscellaneous}

\subsection{\cmd{includesubgraphics}\{\meta{label}\}\{\meta{filename}\}}

Inserts a subfigure (via \pkg{subcaption}) at 90\% text width with a
bold, large label but without a sub-caption. Useful when you want
panel labels (\textbf{A}, \textbf{B}, \ldots) without separate captions.

\subsection{\cmd{protectedemail}\{\meta{address}\}}

Renders an email address in \texttt{typewriter} font inside a TikZ
node. This prevents some simple web scrapers from harvesting the
address as plain text.

%% ---------------------------------------------------------------
\section{Backward Compatibility}

An \texttt{ansgar.sty} shim is distributed alongside \pkg{psychtools}.
It contains a single line:
\begin{verbatim}
  \RequirePackage{psychtools}
\end{verbatim}
Documents that used \cmd{usepackage}\{ansgar\} continue to work
without modification.

%% ---------------------------------------------------------------
\section{License}

\pkg{psychtools} is distributed under the
\href{https://www.latex-project.org/lppl.txt}{LaTeX Project Public
License}, version~1.3c or later.

\end{document}
